---
title: "My R Markdown Primer"
output: html_document
date: "2025-11-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

R Markdown is a way to combine regular text, R code, and formatted math in a single file. 
Instead of writing a report in Word and then running code somewhere else, I can keep 
everything together. When I knit the document, R runs all of the code chunks and 
automatically updates the results, which helps with reproducible analysis.

## What is R Markdown?

R Markdown lets me write in plain text using simple symbols for formatting. 
For example, I can use `#` for headings, `*` for bullet points, and backticks for code. 
In the same document, I can also create R code chunks that are actually executed when 
I knit the file. The final output looks like a polished report, but it is still connected 
to the code that generated the results, which makes it easier to fix mistakes and re-run 
the analysis later.

## Using LaTeX Math

R Markdown also supports LaTeX math, which is useful for writing equations. 
I can write inline math like this: $ \alpha + \beta = \gamma $ inside a sentence.

Here is a displayed equation:

$$
\bar{x} = \frac{1}{n} \sum_{i=1}^{n} x_i
$$

Using LaTeX makes it easier to show more formal math in my reports.

## Loading a Dataset

```{r load-data}
library(ggplot2)
data("mpg")

head(mpg)
```

## Simple Analysis and Plot

```{r simple-plot}
summary(mpg$hwy)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  labs(
    title = "Engine Displacement vs Highway MPG",
    x = "Engine Displacement (L)",
    y = "Highway MPG"
  )
```

## Reflection

Using R Markdown was different at first because it’s not like writing a normal paper. 
But once I understood the Markdown symbols and how the code chunks work, it actually 
made everything easier. I liked that knitting the file automatically updates the plot 
and results without me having to copy/paste anything.

The only issues I ran into were small mistakes, like forgetting a backtick or messing 
up the header. But once I fixed those, it knitted fine. Overall, I like how organized 
R Markdown feels, and I definitely see why people use it for reproducible analysis.